---
title: "DATA1001/1901 Lab12"
subtitle: "Tests for relationships"
author: "© University of Sydney 2020"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

<br>

# Live Demo

We will demonstrate a chi-squared test and regression test.

# Coding Challenge

We will look at the `cats` data set from the `MASS` package. Using the `summary` function find the p-value of the T-test to test the slope of the variables `Hwt` (Heart Weight) against `Bwt` (Body Weight).

<details><summary>Solution</summary>
```{r, warning = F, message=F}
library(MASS)
library(tidyverse)

model = lm(Hwt ~ Bwt, data = cats)

# First check assumptions
plot(model, which = 1:2)


```


The plots show that the assumptions for a linear model are okay. The residuals are homoscedastic and normally distributed.

```{r}
summary(model)   # Gives us the p-value
```

</details>