---
title: "DATA1001/1901 Lab10"
subtitle: "Hypothesis Testing: Proportion Test"
author: "© University of Sydney 2021"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

<br>

# Live Demo

We will give a quick demo of the exercises below.

<br>

#  Probiotic treatment for peanut allergy  

Perform the same test as your Worksheet, in R, using the Sample **Sum**, and then Sample **Mean**.
Check you understand the purpose of each line of code.

```{r,eval=F}
# Write down hypotheses
## Ho:The proportion of patient who respond to the treatment (“desensitise to peanut allergy”) is 0.8
## H1: The proportion of patient who respond to the treatment (“desensitise to peanut allergy”) is more than 0.8.

# H: Set up the box model representing the null hypothesis

## Model: 1= Desensitisation; 0 = Not desensitisation; ie Ho: p=P(Desensitisation) = 0.8
box=c(0,1,1,1,1)    # This represents 20% 0s and 80% 1s

## Calculate the mean and SD of the box
mean(box)
library(multicon)
popsd(box)

# T [for Sample SUM)
## Work out the EV and SE of the Sample Sum
n = 29
ev = n*mean(box)
ev
se = sqrt(n)*popsd(box)
se

## Note the OV of the Sample Sum from the data
ov = 26   # give data

## Calculate the Test Statistic
ts = (ov-ev)/se   
ts

# P: Calculate the p-value using the Normal mode for the Sample Sum (CLT).
1-pnorm(ts)   # 1 sided test, as "more than" in H1
```


```{r,eval=F}
# H: Set up the box model representing the null hypothesis

## Model: 1= Desensitisation; 0 = Not desensitisation; ie Ho: p=P(Desensitisation) = 0.8
box=c(0,1,1,1,1)    # This represents 20% 0s and 80% 1s

## Calculate the mean and SD of the box
mean(box)
library(multicon)
popsd(box)

# T [for Sample MEAN]
## Work out the EV and SE of the Sample Mean
n = 29
ev = mean(box)
ev
se = popsd(box)/sqrt(n)
se

## Note the OV of the Sample Mean from the data
ov = 26/n   # give data

## Calculate the Test Statistic
ts = (ov-ev)/se   
ts

# P: Calculate the p-value using the Normal mode for the Sample Sum (CLT).
1-pnorm(ts)   # 1 sided test, as "more than" in H1
```


<br><br>

# Pepsi and Coke  

Perform the same test as your Worksheet, in R, using the Sample **Sum** and **Mean**.

```{r,eval=F}
# Write down hypotheses
## Ho: The proportion of people who prefer Coke over Pepsi is 0.5.
## H1: The proportion of people who prefer Coke over Pepsi is not 0.5.

# H: Set up the box model representing the null hypothesis

## Model: 1= Coke; 0 = Pepsi; equal proportions, ie Ho: p=P(Coke) = 0.5
box=c(0,1) 

## Calculate the mean and SD of the box
mean(box)
library(multicon)
popsd(box)

# T [for Sample SUM)
## Work out the EV and SE of the Sample Sum
n = 1000
ev = n*mean(box)
ev
se = sqrt(n)*popsd(box)
se

## Note the OV of the Sample Sum from the data
ov = 0.52*n

## Calculate the Test Statistic
ts = (ov-ev)/se   
ts

# P: Calculate the p-value using the Normal mode for the Sample Sum (CLT).
2*(1-pnorm(ts))   # 2 sided, as H1: "Not" 0.5
```

Note: This could have been solved using the sample **Mean**. 

```{r,eval=F}
# H: Set up the box model representing the null hypothesis

## Model: 1= Coke; 0 = Pepsi; equal proportions, ie Ho: p=P(Coke) = 0.5
box=c(0,1) 

## Calculate the mean and SD of the box
mean(box)
library(multicon)
popsd(box)

# T [for Sample MEAN)
## Work out the EV and SE of the Sample Mean
n = 1000
ev = mean(box)
ev
se = popsd(box)/sqrt(n)
se

## Note the OV of the Sample Mean from the data
ov = 0.52

## Calculate the Test Statistic
ts = (ov-ev)/se   
ts

# P: Calculate the p-value using the Normal mode for the Sample Sum (CLT).
2*(1-pnorm(ts))   # 2 sided
```

<br><br>

# Uber 

Perform the same test as your Worksheet, in R, using both the sample **Sum** and then **Mean**.


```{r,eval=F}
# Write down hypotheses
## Ho: The current Sydney customer satisfication with Uber is 0.8
## H1: The current Sydney customer satisfication with Uber is not 0.8

# H: Set up the box model representing the null hypothesis

# Model: 1= Happy with Uber in Sydney; 0 = Unhappy; 80% proportions, ie Ho: p=P(Happy with Uber Sydney) = 0.8
box=c(0,1,1,1,1) 

## Calculate the mean and SD of the box
mean(box)
library(multicon)
popsd(box)

# T [for Sample SUM)
## Work out the EV and SE of the Sample Sum
n= 1200
ev = n*mean(box)
ev
se = sqrt(n)*popsd(box)
se

## Note the OV of the Sample Sum from the data
ov = 948

## Calculate the Test Statistic
ts = (ov-ev)/se   
ts   

# P: Calculate the p-value using the Normal mode for the Sample Sum (CLT).
2*(pnorm(ts))   # 2 sided, as H1 is "not".  # Note TS is negative, hence lower tail area.
```


```{r,eval=F}
# H: Set up the box model representing the null hypothesis

# Model: 1= Happy with Uber in Sydney; 0 = Unhappy; 80% proportions, ie Ho: p=P(Happy with Uber Sydney) = 0.8
box=c(0,1,1,1,1) 

## Calculate the mean and SD of the box
mean(box)
library(multicon)
popsd(box)

# T [for Sample MEAN
## Work out the EV and SE of the Sample Mean
n= 1200
ev = mean(box)
ev
se = popsd(box)/sqrt(n)
se

## Note the OV of the Sample Mean from the data
ov = 948/n

## Calculate the Test Statistic
ts = (ov-ev)/se   
ts   

# P: Calculate the p-value using the Normal mode for the Sample Sum (CLT).
2*(pnorm(ts))   # 2 sided  # Note TS is negative.
```


