---
title: "DATA1001 Lab2"
subtitle: "Data & Graphical Summaries"
author: "© University of Sydney 2020"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---

# Live Demo

1. We will demonstrate how to classify variables and isolate a variable. This is called Initial Data Analysis (IDA).

```{r}
# lists the top 6 rows of the data
head(iris)

#lists the top 2 rows
head(iris,2)

# lists the number of rows & number of columns
dim(iris)

#lists the names of the variables
names(iris)

# classify variables
str(iris)

# isolate a variable
iris$Sepal.Length
```


Note: Keep in mind that R has slightly different terms for types of variables.

“factor” = qualitative

“numeric” = quantitative


2. We will demonstrate how to produce a histogram and a barplot in base R.

```{r}
# produce a histogram (quantitative variable)
hist(iris$Sepal.Length)

# produce a barplot (qualitative variable)
barplot(table(iris$Species))
```

3. We will demonstrate how to produce a histogram and a barplot in ggplot.

```{r}
# install.packages("ggplot2")  # You will need to run this line of code if you haven't installed ggplot2 before
library(ggplot2)

# produce a histogram (quantitative variable)
p1 = ggplot(iris, aes(x = Sepal.Length))  # Defines the x axis (1 variable).
p1 + geom_histogram() # Represents the data by histogram. 

# produce a barplot (qualitative variable)
p2 = ggplot(iris, aes(x=Species))
p2 + geom_bar()
```


# Have a go

1. List the bottom 2 rows of the iris data.

```{r}
# Write your code in here
```

2. Select the `Sepal.Width` variable and produce a histogram in base R, and then ggplot.

```{r}
# Write your code in here
```

3. Select the `Species` variable and produce a barplot, shaded green, in both base R and then ggplot.

```{r}
# Write your code in here
```


4. Experiment yourself.

# Coding Challenge

Using the `iris` data create a histogram of the `Petal.Length` variable using `ggplot` and the `geom_histogram` functions. Set the number of bins or columns in the plot to be 14 using the argument `bins=14` argument inside the `geom_histogram` function.

What is the height of the largest column in the resulting plot?


<details><summary>Solution</summary>
```{r}
library(ggplot2)
ggplot(iris, aes(x=Petal.Length)) + geom_histogram(bins = 14)
```

Answer = 35
</details>
