---
title: "DATA1901 Lab2 (Ext)"
subtitle: "Design of Experiments"
author: "University of Sydney"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
    
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>


# Examining Simpson's paradox with the Smoking Study (UK)
Here we examine the Smoking data from Week 1 lectures. 


## Upload the data

Upload the data direct from a Canvas website url. 

```{r}
smoking = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/2018S2/data/simpsonssmoking.csv",header=T)
```


## IDA
- What is the size of the data file? What do the rows and columns represent? Is this the full data from the UK study, or a summary?

```{r}
dim(smoking)
names(smoking)
```

- Can you see any patterns?

```{r}
smoking
```

## Research Question: Is the mortality rate higher for smokers or non-smokers?

### 1st, consider the overall mortality rates 

- Calculate the mortality rate for smokers.

```{r}
sum(smoking$SmokersDied)/sum(smoking$SmokersDied+smoking$SmokersSurvived)
```

- Calculate the mortality rate for non-smokers.

```{r}
# Put your code here
```

- Create a barplot of the 2 mortality rates.

```{r}
barplot(c(23.9,31.4),names.arg=c("Smoker","Non-smoker"),col=c("lightgreen","lightblue"),main="Mortality Rates for Women over 20 years")
```

- What do these results seem to imply?

### 2nd, examine the mortality rate by age groups 
- Did more smokers or non-smokers die in the 18-24 age group?

```{r}
# Consider Smokers 18-24
sum(smoking$SmokersDied[1])/sum(smoking$SmokersDied[1]+smoking$SmokersSurvived[1])
```

```{r}
# Consider Non-Smokers 18-24
sum(smoking$NonSmokersDied[1])/sum(smoking$NonSmokersDied[1]+smoking$NonSmokersSurvived[1])
```

Note: `smoking$SmokersDied[1]` selects the 1st entry of smoking$SmokersDied.

- Did more smokers or non-smokers die in the 25-34 age group?

```{r}
## Put your code here
```

### Simpson's Paradox
Explain Simpson’s Paradox, with reference to the UK Smoking data.

# Using kable and ggplot

- Note, that you can format the table using the following package and command.

```{r}
# You can format the table by using kable.
# install.packages("knitr")
library(knitr)
kable(smoking)
```

- Compare all age groups, to see what pattern emerges. Are there anomalies? Postulate a reason.

- Run the following code. You will learn how to use `ggplot` next week!

```{r}
library(ggplot2)
age_groups <- data.frame(smoker = factor(c(rep("Yes", 7), rep("No", 7))), age_group = factor(c("18-24", "25-34", "35-44", "45-54", "55-64", "65-74", "75+")), perc_dead = c(3.6, 2.4, 12.8, 20.7, 44.3, 80.5, 100, 1.6, 3.2, 5.8, 1.3, 33.1, 78.3, 100))
plt2 = ggplot(age_groups, aes(age_group, perc_dead)) + geom_bar(aes(fill = smoker), 
    stat = "identity", position = "dodge") + scale_fill_manual(values = c("lightblue", 
    "lightgreen"))
plt2 + labs(x = "Age Group", y = "Mortality Rate (% Died)") + ggtitle("Mortality Rates for Women over 20 years, by age group")
```

