---
title: "DATA1001/1901 Lab3 Summary Sheet"
subtitle: "Graphical Summaries"
author: "© University of Sydney 2020"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fa fa-th-list"></i> Summary {-}

You are beginning to learn how to:

- pose research questions
- choose what graphical summaries are appropriate for given variables
- interpret the output.

Below is an exemplar of lots of research questions applied to the Road Fatailities dataset.

<br>

Type of Variable | Referred to in R |
-------- | ------- | 
Qualitative / Categorical | `factor`
Quantitative / Numerical | `num`
| |

Type of Data | Type of Graphical Summary | In R |
------------------------ | -------------------------- | -------------- | 
1 Qualitative Variable | Barplot | `barplot()` |
2 Qualitative Variables |  Double (clustered) Barplot |  `barplot()` |
1 Quantitative Variable | Histogram or Boxplot  | `hist()`, `boxplot()`|
2 Quantitative Variables | Scatterplot | `plot()`|
1 Quantitative & 1 Qualitative Variable | Double (comparative) boxplot | `boxplot()`|
| | |

<br>


# <i class="fas fa-car"></i> Australian Road Fatalities

Consider the Australian road fatalities from 1989 (a bigger version of the data used in Week 2 lectures). The data is sourced from [BITRE](https://bitre.gov.au/statistics/safety/fatal_road_crash_database.aspx).

**How to proceed?**

- Read through each question.
- Try to guess the code, and then check the given code.

<br>

## Initial Data Analysis

- Upload the data.

```{r}
# Read data from url into R
road = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/data/AllFatalities.csv")
```

Note: An alternative way is to download the data from Canvas, store the data in `DATA1001files/data` and upload from there. You will need to use this method in future projects, when you upload your own data.

```{r,eval=F}
# Read data from url into R
road = read.csv("data/AllFatalities.csv",header=T)
```

- Produce a snapshot of the data. 
```{r,eval=F}
str(road)
```


<br>

## Research Questions

### Were there more fatalities on a certain day of the week? 

- Here we consider **1 qualitative variable**: the road fatalities across the *days of the week*.

- 1st isolate the variable Dayweek. Check how R classifies it. Produce a barplot. What is annoying about it?

```{r}
class(road$Dayweek)
barplot(table(road$Dayweek))
```

- We can re-order the categories for `dayweek` and the produce a barplot. What pattern emerges? Suggest possible reasons for it? 

```{r}
orderdayweek = ordered(road$Dayweek, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
barplot(table(orderdayweek))
barplot(table(orderdayweek),las=2)
```

<br>

### Was there any pattern in how buses were involved in fatalities, on different days of the week?

- Here we consider **2 qualitative variables**: the road fatalities across the *days of the week*, cross-classified by *bus involvement*.

- Is there any pattern?

```{r}
road1 = table(road$Bus_Involvement, road$Dayweek)
road1
barplot(road1, main = "Fatalities by Day of the Week and Bus Involvement", xlab = "Day of the week", 
    col = c("lightblue", "lightgreen"), legend = rownames(road1))
```


<br>

### Was there any pattern in how heavy rigid trucks were involved in fatalities, on different days of the week?

- Here we consider **2 qualitative variables**: the road fatalities across the *days of the week*, cross-classified by *heavy rigid truck involvement*.

- Investigate whether the involvement of heavy rigid trucks differs across the days?

```{r}
road2 = table(road$Hvy_Rigid_Truck_Involvement, road$Dayweek)
road2
barplot(road2, main = "Fatalities by Day of the Week and Heavy Rigid Involvement", xlab = "Day of the week", 
    col = c("lightpink","lightblue", "lightgreen"), legend = rownames(road2))
```

<br>

### Were there more fatalities in certain age groups?

- Here we consider **1 quantitative variable**: *fatalities*.

- 1st isolate the variable Age. How does R classify it?

```{r}
class(road$Age)
```

- Change the classification to a quantitative variable.
```{r}
road$Age = as.numeric(as.character(road$Age))
class(road$Age)
```

- Produce graphical summaries. What patterns are revealed?
```{r}
hist(road$Age, prob=T)
boxplot(road$Age)
```

- We can customise the plots.
```{r}
hist(road$Age,freq=FALSE,main="Histogram",ylab="Probabilities", col="green")
boxplot(road$Age,horizontal=TRUE,col="red")
```

<br>

### Does biological gender affect the number of fatalities across age groups?

- Here we consider **1 quantative variable divided by 1 qualitative variable**.

- Control for biological gender - ie consider fatalities by age divided by biological gender.
```{r}
ageF = road$Age[road$Gender == "Female"]
ageM = road$Age[road$Gender == "Male"]
par(mfrow = c(2, 1))
boxplot(ageF,horizontal=T, col="light blue")
boxplot(ageM,horizontal=T)
```

- You can put 2 plots next to each other.
```{r}
par(mfrow=c(2,1))
boxplot(ageF,horizontal=T, col="light blue")
boxplot(ageM,horizontal=T)
```

<br>

