---
title: "DATA1001/1901 Lab4 Worksheet (Answers)"
subtitle: "Normal Model"
author: "University of Sydney"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>


# <i class="fa fa-crosshairs"></i> Aim {-}

- Now we will see how easy it is in R to find areas under the Normal curve and percentiles.

- You should still draw a quick sketch of the Normal curve for each section, highlighting the relevant area, so you can visualise the problem, before you use R!

<br>

# Standard Normal Curve

The Standard Normal curve has mean 0 and standard deviation (SD) 1, and is written as  N(0,1)

- What is the area under the curve between -1 and 1?

```{r}
pnorm(1)-pnorm(-1)
```

- What is the  area between -2.5 and 2.5.

```{r}
# Write your code here
```


- What is the 84% percentile?

```{r}
qnorm(0.84)
```

- What is the 60% percentile.

```{r}
# Write your code here
```

<br>


# General Normal Curve

A General Normal curve has mean 10 and standard deviation (SD) 3, and is written as  N(10,3^2):

- What is the area under the curve lower than 7?

```{r}
pnorm(7,10,3) 
```

- What is the 30% percentile?

```{r}
# Write your code here
```

<br>

# <i class="fas fa-football-ball"></i> Australian men's AFL team

Using the [ABS](http://www.abs.gov.au/ausstats/abs@.nsf/Lookup/4338.0main+features212011-13 ), assume that the height of Australian mean can be modelled by a Normal curve, with mean = 175.6cm and SD = 10cm.

- What percentage of Australian men are between 165.6cm and 185.6cm?

```{r}
pnorm(185.6,175.6,10) - pnorm(165.6,175.6,10)
```


- If 50% of Australian men are below a certain height, what is that height?

```{r}
qnorm(0.5,175.6,10)
```


<br>

In the Australian Football League (AFL) recruiters tend to look for tall male players. According to the [HeraldSun](http://www.heraldsun.com.au/sport/afl/size-matters-at-afl-level/story-e6frf9jf-1226650225771), the minimum height that recruiters look for is 180cm. Using the Normal curve above, estimate the chance of finding an Australian man able to be recruited for AFL?

```{r}
# Write your code here
```

<br>

West Coast ruckman Nick Naitanui is 201cm tall. What is the chance of finding an Australian male taller than Nick?

```{r}
# Write your code here
```

