---
title: "DATA1001/1901 Lab4"
subtitle: "Normal Model"
author: "© University of Sydney 2021"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>


# Live Demo

We will demonstrate how to:

- add a Normal curve to a histogram
- produce Normal probabilities (pnorm) and Normal quantiles (qnorm).

## Adding normal curve to histogram {-}
### In base R {-}

```{r}
# How to add a Normal curve to a histogram
hist(iris$Petal.Length, freq=F)
curve(dnorm(x, mean = mean(iris$Petal.Length), sd = sd(iris$Petal.Length)), col = "lightblue", lwd = 6, add = TRUE)
```

### In ggplot {-}

To produce the same plot in `ggplot` we can utilise the `stat_function` geometry which draws a function over the plot.

```{r}
library(ggplot2)

m = mean(iris$Petal.Length)
s = sd(iris$Petal.Length)

p1 = ggplot(iris, aes(x = Petal.Length)) + 
  geom_histogram(aes(y = ..density..)) +
  stat_function(fun = dnorm, args = list(mean = m, sd = s), colour="lightblue", size=2)
p1
```


```{r}
# Produce Normal probabilities
pnorm(7, mean = mean(iris$Petal.Length), sd = sd(iris$Petal.Length))
```


```{r}
# Produce Normal quantiles
qnorm(0.5, mean = mean(iris$Petal.Length), sd = sd(iris$Petal.Length))
```

# Coding Challenge

Create a normal model for the `Sepal.Length` variable from the `iris` data set and use this model to answer the following question. What percentage of the sample (to the nearest percent) would you expect to have a sepal length between 5 and 6.

_Hint: You will need to use the `pnorm` function twice_ 


<details><summary>Solution</summary>
```{r}
m = mean(iris$Sepal.Length)
s = sd(iris$Sepal.Length)
pnorm(6, m, s) - pnorm(5, m, s)

```
Answer = 42%
</details>

