---
title: "DATA1001/1901 Lab6"
subtitle: "Linear Model"
author: "© University of Sydney 2021"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---

<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# Live Demo

We will demonstrate some non-linear modelling on iris, and then do a quick demo of the rest of the exercises below.

```{r}
# Consider the iris data
```


<br>

# Blood Alcohol Concentration (BAC) 

Blood Alcohol Concentration (BAC) is the percentage of alcohol in the blood. Referring to this [Drinkwise article](https://drinkwise.org.au/drinking-and-you/is-there-such-a-thing-as-safe-drink-driving/#):

- What is considered a ["standard drink"](https://drinkwise.org.au/drinking-and-you/how-much-have-you-had-to-drink/#) in Australia?
- What is the legal BAC for [driving a car](https://drinkwise.org.au/drinking-and-you/is-there-such-a-thing-as-safe-drink-driving/#) in Australia (learner, provisional,full)?
- What are the possible effects of alcohol on driving?

The data in `AlcoholConcentration` contains the blood concentration (in ng/mL) at time t hours after consuming alcohol.

```{r}
bac = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/data/AlcoholConcentration.csv")
names(bac)[1] = "TimeHours"     # Use this line if the csv doesn't load the correct column name
```

## Executive Summary

## Initial Data Analysis (IDA)

```{r,results='hide', fig.show = 'hide'}
str(bac)
```

## Research Question:  Is there a relationship between BAC and time?

Here we consider 2 quantitative variables: `Time` and `Conc`.

- Produce a scatterplot.

```{r,results='hide', fig.show = 'hide'}
# In base R
Time = bac$TimeHours
Conc = bac$BloodAlcoholConcentration
plot(Time, Conc, xlab="Time (hr)", ylab="Conc (units)", pch=16, cex.lab=1.3, col="red")
```

```{r, results='hide', fig.show='hide', warning=F}
# In ggplot
library(ggplot2)

scatter_plot = ggplot(bac, aes(x = TimeHours, y = BloodAlcoholConcentration)) +
  geom_point(colour = "red", size = 2) +
  xlab("Time (hr)") + 
  ylab("Conc (units)")
scatter_plot
```
    

- What do you observe?

### Fit the following linear model to the data.

\[ BAC = 0.100 - 0.015 t    \]

```{r,results='hide', fig.show = 'hide'}
# In base R

plot(Time, Conc, xlab="Time (hr)", ylab="Conc (units)", pch=16, cex.lab=1.3, col="red")
abline(a=0.100,b=-0.015, col = "darkgreen", lwd = 3)
# Plot a horizontal line at 0.05
abline(h=0.05, col = "purple", lwd = 1, lty=2)
```

```{r,results='hide', fig.show = 'hide'}
# In ggplot

# We have saved the plot from before so we can add the linear model line to it
scatter_plot + geom_abline(intercept = 0.1, slope = -0.015, colour = "darkgreen", size = 1.2) + 
  geom_hline(yintercept = 0.05, linetype = "dashed", colour = "purple")
  
```

- From the plot, how long does it take to be legal to drive?

- Using the scatter plot and residual plots, assess the suitability of the model.

```{r,results='hide', fig.show = 'hide'}
par(mfrow=c(1,2))
# Scatter plot
plot(Time, Conc, xlab="Time (hr)", ylab="Conc (units)", pch=16, cex.lab=1.3, col="red")
abline(a=0.100,b=-0.015, col = "darkgreen", lwd = 3)
# Residual plot
predictedvalues = 0.100-0.015*Time
residuals = Conc - predictedvalues
plot(Time, residuals, ylab = "residuals")
abline(h = 0, col="blue")
```

## Further research

What does this [report](https://www.aph.gov.au/DocumentStore.ashx?id=c1bbe04c-e3ad-442e-93d9-9d99c639b7b9&subId=411549) find about BAC in Australian night time entertainment districts? How was linear regression used?

<br>

# Blood Ice Concentration (BIC) 

Methamphetamine (or ice/speed/meth) is a highly addictive central nervous system (CNS) stimulant that is used as a recreational drug. 

The data in `MethConcentration` contains the blood concentration (in ng/mL) at time $t$ hours after using ice.

```{r}
bic = read.csv("http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/data/MethConcentration.csv")
names(bic)[1] = "TimeHours"     # Use this line if the csv doesn't load the correct column name
```

## Domain knowledge

From the following [ADF article](https://adf.org.au/drug-facts/ice/):

- What are the different forms of meth?
- What are possible side-effects of using crystal meth (ice)?
- How long can the ice 'high' last?
- What are some medical uses of meth?
- Why does Australia now regulate the sale of pseudoephedrine (a common nasal decongestant)?

## Initial Data Analysis (IDA)

```{r,results='hide', fig.show = 'hide'}
str(bic)
```

## Research Question: Is there a relationship between BIC and time?

Here we consider 2 quantitative variables: `Time` and `Conc`.

- Produce a scatterplot.

```{r,results='hide', fig.show = 'hide'}
Time =  bic$TimeHours
Conc = bic $BloodConcentration
# Insert your own code
```

- What do you observe?

### Fit a linear model: $y = a + bt$

```{r,results='hide', fig.show = 'hide'}
# Insert your own code
```

- What is the correlation coefficient? 

```{r,results='hide', fig.show = 'hide'}
# Insert your own code
```

- Check the residual plot. What is the problem?

```{r,results='hide', fig.show = 'hide'}
# Insert your own code
```

Note: If the linear model is `L = lm(y ~ x)`, then the residuals are found in `L$residuals` or `resid(L)`, and the residual plot is `plot(x,L$residuals)`.


### Fit a quadratic model: $y = a + bt + c t^2$

```{r,results='hide', fig.show = 'hide'}
# Quadratic model
Time2 = Time^2
quad.model = lm(Conc ~ Time + Time2)
summary(quad.model)
# Predicted values
set.seed(1)
timevalues = seq(0, 50, 0.1)
predictedvalues = predict(quad.model, list(Time = timevalues, Time2 = timevalues^2))
# Plot
plot(Time, Conc, pch = 16, xlab = "Time (s)", ylab = "Conc", cex.lab = 1.3, col = "indianred")
lines(timevalues, predictedvalues, col = "darkgreen", lwd = 3)
```

### Try $y = 47e^{−0.06t}$, for $t > 3$.

```{r,results='hide', fig.show = 'hide'}
plot(Time,Conc)
curve(47*exp(-0.06*x),3,50,add=T)
```

For full code for fitting an exponential, see your [lecture notes](http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/lectures/18.NonLModels.html).

### Try your own choice.

```{r,results='hide', fig.show = 'hide'}
# Insert your own code
```

<br>

# Coding Challenge

Using the BAC data set, fit a polynomial model to the data of order 3. Use `BloodAlcoholConcentration` as your outcome variable (y) and `TimeHours` as the predictor (x). Next calculate the residuals of the fitted model to the data and plot them using `ggplot`. How many of the residuals are greater than 0.01?

<details><summary>Solution</summary>
```{r, message=F, warning=F}
library(ggplot2)

# Produce the model
poly_model = lm(BloodAlcoholConcentration ~ poly(TimeHours, 3), data = bac)

# Add the residuals to the bac data frame
bac$predicted_values = predict(poly_model)
bac$residuals = bac$BloodAlcoholConcentration - bac$predicted_values

# Plot the residuals using ggplot and add a line at y = 0.01 to make it easy to see how many are above this line.
ggplot(bac, aes(x=TimeHours, y = residuals)) + geom_point() + geom_line(aes(y = 0.01), colour="red")

# Alternate solution without plotting
sum(bac$residuals > 0.01)

```
</details>

# DATA1901 Extension

## Windmark's Formula

[Windmark's Formula](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4361698/) states that:
\[ BAC = \frac{100A}{rW} − V t \]
where

- A = amount of alcohol consumed (in grams)
- r = Windmark's factor, measuring the % of body mass that is water (women \approx 0.6; men \approx 0.7)
- W = body mass (in grams)
- V = rate at which the boy eliminates alcohol (eg V=0.015%)
- t = time since commencing drinking (in hours)

Assume:

- the average Australian male [weighs 85.9kg](http://www.abs.gov.au/ausstats/abs@.nsf/Lookup/4338.0main+features212011-13) and drinks 2.5 beers in an hour. 
- a beer contains 1.5 standard drinks, where a standard drink is 10g alcohol. 
- a male drinks 4 beers in a short period of time.

- Show the formula (to 3dp) for BAC is 
\[ BAC = 0.100 - 0.015 t    \]

- What is his BAC after 1 hour?

- How long does it take for him to be legal to drive?

More Reading:
[BAC Formulae](http://www.kenderdinemathstutoring.com.au/downloads/3011085/BAC+formula+questioned+amended.pdf)
[Modelling drugs](http://www.appstate.edu/~spruntwh/bac_ncctm.pdf)







