---
title: "DATA1001 Lab7 Worksheet"
subtitle: "Understanding Chance"
author: "© University of Sydney 2020"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fa fa-crosshairs"></i> Aim {-}

At the end of this Worksheet, you should be able to:

- model an appropriate scenario by the Binomial model
- explain the Prosecutor's fallacy

<br>

**Binomial Formula:**

- See formula in lectures [here](http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/lectures/21.BinomialFormula.html#21).

- Suppose we have n independent, binary trials, with P(event)=p at every trial, and n is fixed.

- The chance that exactly x events occur is:
\[ { n \choose x} p^{x} (1-p)^{n-x} \]


**Hints:** 

- Check you have a sequence of n repeated independent trials.
- Identify the event which is considered a "success", and the constant "success" probability p. 
- Identify the number of successes being considered x. 
- Then use the Binnomial formula. 

<br>

# Scenarios

## Call centre

Here's the details:

- A company "cold calls" their current customers to try to get them to upgrade to a new phone plan.
- An employee completes on average 5 calls per hour, with success rate 40%.
- The employee gets a \$50 bonus for getting all 5 customers to upgrade.
- What is the chance that the employee gets a bonus?

Assumptions:

- Let each phone call the employee makes be an independent trial, with success rate 40%.  So p = 0.4.
- The employee makes 5 calls per hour. So n = 5.
- We want to consider x = 5.

Model:

- Let X = number of successful phones calls (ie customer upgrades) out of 5 ~ Bin(n=5, p=0.4)
- The chance that x = 5 is:

\[ { 5 \choose 5} (0.4)^{5} (0.6)^{0} = 0.01024 \]

Conclusion: In an hour, there's a 1% chance that the employee gets the bonus!

## Have a go

Here's the details:

- Suppose a new diagnostic test can detect a certain virus with 90% accuracy.
- 1000 people are tested for the virus, using the new diagnostic test.
- What is the chance that exactly 9% of the people get an *inaccurate* result?

Assumptions:

- Let each person having the medical test be an independent trial, with success rate 90%.  So p = 0.9.
- 1000 people are tested. So n = 1000.
- We want to consider x = 910 (91% of 1000 are accurate, as 9% are inaccurate).

Model:

- Let X = number of successful medical detections of the virus in 1000 trials ~ Bin(n=1000, p=0.9)
- The chance that x = 910 is:

\[ { 1000 \choose 910} (0.9)^{910} (0.1)^{90} =  \]

```{r}
choose(1000,910)*(0.9)^910*(0.1)*90
```

Conclusion: The chance of getting exactly 9% inaccurate tests is tiny.

<br>

Note: More likely we would be asking "What is the chance that less than 9% of the people get an *inaccurate* result?" which is the same as "What is the chance that 91% or more of the people get an *accurate* result?"

So the answer is:

```{r}
# Summing up all the cases 910-1000
x=c(910:1000)
sum(choose(1000,x)*(0.9)^x*(0.1)^(1000-x))

# Using dbinom
x=c(910:1000)
sum(dbinom(x,1000,0.9))

#  Using pbinom
1-pbinom(909,1000,0.9)   # 910 or more successes, is 1- 909 or less successes.
```

Conclusion: The chance of getting less than 9% inaccurate tests is about 16% (to 2dp).

<br>

# <i class="fas fa-gavel"></i> Prosecutor's Fallacy

What is the prosecutor's fallacy? What was the mistake in statistical thinking in the Sally Clarke trial? 

<br>

# <i class="fab fa-spotify"></i> Spotify  

Spotify is an online music streaming service, which allows the user to create and share play lists.

## Song list
- Create a potential Spotify list, by listing your top 10 favourite songs. 

For example:

|Song|Name|Artist|Australian Artist (Yes/No)|
|---|---|---|---|
|1|Skinny Love|Birdy|No|
|2|Better Together|Jack Johnson|No|
|---||||
|10||||
|||||

- In how many different ways could Spotify shuffle this list?

## Play 2 Songs (with replacement)
Suppose Spotify randomly plays 2 songs, with replacement.

- How many different play lists are there?

- What is the chance of getting a certain song twice?

- What is the chance of getting any song twice? Simulate this situation, using 1000 replications.

```{r, results='hide'}
set.seed(2)
songs=c(1:10)
r=replicate(10000,sample(songs,2,replace=T))
length(which(r[1,]==r[2,]))/10000
```

- What is the chance of getting a certain song twice, given it was the 1st song? 

Check your answers:
```{r, results='hide'}
# number of ways to shuffle 10 songs (with replacement)
factorial(10)
# total number of playlists
10*10
# chance of getting a certain song twice
1/(10*10)
# chance of getting any song twice
10/(10*10)
# chance of not getting any song twice
1-(10/(10*10))
```

## Play 2 Songs (without replacement)
Now suppose Spotify randomly plays 2 songs, without
replacement.

- How many different play lists are there?

- What is the chance of getting your favourite play list?

- What is the chance of getting a certain song twice? Simulate this situation.

```{r, results='hide'}
set.seed(2)
songs=c(1:10)
r=replicate(10000,sample(songs,2,replace=F))
length(which(r[1,]==r[2,]))/10000
```

Check your answers:
```{r, results='hide'}
# total number of playlists (w/o replacement)
10*9
# chance of getting your favourite playlist
1/(10*9)
# chance of getting a certain song twice
0
# chance of not getting any song twice
1-(10/(10*10))
```

- According to this [article](https://www.lifewire.com/is-itunes-shuffle-random-1999249), how does ipod shuffle it's playlists?

Extension: 
<i class="fa fa-link"></i> [Spotify and Data](https://www.safaribooksonline.com/library/view/strata-data-conference/9781491976326/video314445.html?utm_source=oreilly&utm_medium=newsite&utm_campaign=strata-ny-17-keynote-highlight-music-the-window-into-your-soul)

## Properties

- Which of the previous events are independent, dependent or mutually exclusive?

## Modelling by Binomial

Suppose that there are 5% of Australian artists on Spotify. You randomly sample 10 songs, with replacement.

- How could this situation be modelled by throwing a coin?

- What is the probability of sampling 2 Australian artists?

```{r, results='hide'}
# Using Binomial formula
choose(10,2)*(0.05)^2*(0.95)^8
# Speedy way in R
dbinom(2,10,0.05)
```

- What is the probability of sampling 2 or less Australian artists?

```{r, results='hide'}
# Using Binomial formula
dbinom(0,10,0.05) + dbinom(1,10,0.05) + dbinom(2,10,0.05)
# Speedy way in R
pbinom(2,10,0.05)
```

