---
title: "DATA1001/1901 Lab7"
subtitle: "Understanding Chance"
author: "© University of Sydney 2020"
output:
  html_document:
    theme: flatly
    number_sections: yes
    toc: true
    toc_float: true
    toc_depth: 4
    code_folding: show
---

<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>


# Live Demo

We will demonstrate simulation of simple scenarios, including a simple game.

## Simulation

```{r}
set.seed(1)  # This ensures the same simulation, "1" can be whatever you choose

# Simulate 1 fair coin toss
sample(c("H", "T"), 1, replace = T)

# Simulate 1 biased coin toss
sample(c("H", "T"), 1, replace = T,prob = c(0.9, 0.1))

# Simulate 10 fair coin tosses, and summarise chances in a table.

## Option1
sample1 = sample(c("H", "T"), 10,replace=T)
table(sample1)/10

## Option 2
sample2 = replicate(10, sample(c("H", "T"),1,replace=T))
table(sample2)/10

## Option 3
sample3 = rbinom(n = 10, 1, 0.5)
table(sample3)/10
```

## Using binomial function

Suppose we have 10 trials with P(event) = 0.3.

```{r}
# The chance of getting 4 events is:
dbinom(4,10,0.3)  # dbinom(x,n,p)

# The chance of getting less than or equal to 5 events is:
pbinom(5,10,0.3)    # pbinom(x,n,p)
```

<br>

# Game

Here's the rules of a game:

- A game costs $50 to play. 
- You get to throw a fair dice 20 times.
- You get \$8 for any 6s thrown.

Simulate your net win/loss.

```{r}
# Simulate the 20 throws of the fair dice
set.seed(100)
sample3 = sample(c(1:6), 20, replace = T)
table(sample3)
barplot(table(sample3), xlab="dice face", ylab="chance", col= "lightblue")
```

Conclusion: Based on this simulation, your net loss would be \$8*5-\$50 = -\$10.


<br><br>

# Have a go

Here's the rules of another game:

- A game costs $10 to play. 
- You get to throw a biased coin 5 times, where the probability of getting a head is 0.2. 
- You get $20 for each Head you thow. 

Simulate your net win/loss.

```{r}
# Simulate the 5 throws of the biased dice
```


