---
title: "DATA1001 Lab8 Worksheet"
subtitle: "Chance Variability (The Box Model)"
author: "© University of Sydney 2020"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

<br>

# <i class="fa fa-crosshairs"></i> Aim {-}

At the end of this Worksheet, you should be able to draw a "box model" for a given random experiment, and calculate the EV, SE and Normal probabilities for the Sum and Mean of the draws.

<br>

**Recap on the "Box model"**

- See [Lecture23](http://www.maths.usyd.edu.au/u/UG/JM/DATA1001/r/current/lectures/23.TheBoxModel.html#1).

- The "box" represents the population, and the $n$ draws represents thes sample.

- Unless otherwise stated, we assume the draws are made with replacement from the box.

- Take a box with a certain mean and SD. Take $n$ draws to form a sample. The following results follow:

|Focus|Expected Value (EV)|Standard Error (SE)|
|------------|------------|--------------|
|Sum of the Sample| $n \times$ Mean of the Box | $\sqrt{n} \times$ SD of the Box |
|Mean of the Sample| Mean of the Box  | SD of the Box / $\sqrt{n}$|

- If we prefer symbols, we can call the mean and SD of the box $\mu$ and $\sigma$ respectively, then we can write:

|Focus|Expected Value (EV)|Standard Error (SE)|
|------------|------------|--------------|
|Sum of the Sample| $n \mu$|$\sqrt{n} \sigma$ |
|Mean of the Sample|$\mu$|$\sigma$ / $\sqrt{n}$|

<br>

**Hints:** 

- Draw the scaffold for the box model, and fill in the details.
- Check whether you are being asked about the **Sum** OR **Mean** of the draws.
- Use the results above.

<br>
**When you finish:** Upload your answers to the Lab10 link on [Ed](https://edstem.org/courses/3894/discussion/).

<br><br>

# Have a go

Explain the Box Model by an annotated diagram, and example.

<br><br>

# Scenarios

## Box1

What we know:

- A box contains 1,2,3. This is the population.
- We draw 25 times from the box, with replacement. This forms the sample.
- We are interested in the properties of the **Sum** of the 25 draws. 

<br>

(i) Draw the box model, by hand.

<br><br><br>


(ii) Use the box model.

The **Sum** of the sample has:

- EV = $n \times$ Mean of the Box = 25 x mean = 
- SE =$\sqrt{n} \times$ SD of the Box  = 5 x SD = 

(iii) Simulate 2 samples, and see how the 2 **Sums** compare to the EV.

```{r}
# sample 1
set.seed(5)
box=c(1,2,3)
sample = sample(box, 25, replace = T)
table(sample)
sum(sample)

# sample 2
set.seed(10)
box=c(1,2,3)
sample = sample(box, 25, replace = T)
table(sample)
sum(sample)
```


(iv) Now replicate 10000 samples, and see how the histogram of **Sums** matches up with the EV and SE.

```{r}
# replicate 10000 samples, and calculate 10000 sums
set.seed(10)
box=c(1,2,3)
totals = replicate(10000, sum(sample(box, 25, replace = T)))
hist(totals)
## compare the simulation mean and SD to the EV and SE of the model.
mean(totals)
sd(totals)
```

<br>

## Box2

- A box contains 1,2,3,4,5,6. This is the population.
- We draw 16 times from the box, with replacement. This forms the sample.
- We are interested in the properties of the **Mean** of the 16 draws. 


<br>

(i) Draw the box model, by hand.

<br><br><br>

(ii) Use the box model.

The **Mean** of the sample has:

- EV = Mean of the Box = 
- SE = SD of the Box / $\sqrt{n}$   = 

(iii) Simulate 2 samples, and see how the 2 **Means** compare to the EV.

```{r}
# sample 1
set.seed(5)
box=c(1,2,3,4,5,6)
sample = sample(box, 16, replace = T)
table(sample)
mean(sample)

# sample 2
set.seed(10)
sample = sample(box, 16, replace = T)
table(sample)
mean(sample)
```


(iv) Now replicate 10000 samples, and see how the histogram of **Means** matches up with the EV and SE.

```{r}
# replicate 1000 samples, and calculate 1000 sums
set.seed(10)
box=c(1,2,3,4,5,6)
totals = replicate(1000, mean(sample(box, 16, replace = T)))
hist(totals)
## compare the simulation mean and SD to the EV and SE of the model.
mean(totals)
sd(totals)
```


<br>

## Have a go

What we know:

- A box contains 1,3,6. This is the population.
- We draw 25 times from the box, with replacement. This forms the sample.
- We are interested in the properties of the **Sum** of the 25 draws. 

<br>

(i) Draw the box model, by hand.

<br><br><br>

(ii) Use the box model.

The **Sum** of the sample has:

- EV = 
- SE = 

(iii) Simulate 2 samples, and see how the 2 **Sums** compare to the EV.

```{r}
```

(iv) Now replicate 10000 samples, and see how the histogram of **Sums** matches up with the EV and SE.

```{r}
# replicate 10000 samples, and calculate 10000 sums
set.seed(10)
box=c(1,3,6)
totals = replicate(10000, sum(sample(box, 25, replace = T)))
hist(totals)
## compare the simulation mean and SD to the EV and SE of the model.
mean(totals)
sd(totals)
```

<br>

# <i class="far fa-sun"></i> Climate change

Suppose you wanted to conduct a survey on the opinion of Sydney students on climate change.

- Propose a question.
- What is the population and sample?
- What is the parameter of interest?
- What collection method would you use and why?
- What are possible sources of bias?
- Draw the box model.


